Smooth and Persistent Forecasts of German GDP\\ Balancing Accuracy and Stability\thanks{We would like Barbara Rossi and the participants at the Vienna Forecasting Workshop 2025 for their comments and suggestions.}

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{1},...,x_{t})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$. For simplicity, we assume stationarity of all time series involved. In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, representing the (log differenced) GDP, for which $\mathbf{x}_t$ is effectively observed. Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}$, where $L\leq t$.

@@

We can now extend the SSA criterion \ref{crit1} to a multivariate setting. Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the four indicators as well as to GDP. %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
%For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.

>>>>

Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{t},...,x_{1})'$. Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$. For simplicity, we assume stationarity of all time series involved. In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, representing the (log differenced) GDP, for which $\mathbf{x}_t$ is effectively observed. Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}$, where $L\leq t$.

@@

We can now extend the SSA criterion \ref{crit1} to a multivariate setting. Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$. Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the four indicators as well as of GDP. %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
%For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.
